---
title: "Reproducible Data Project 1"
author: "Rick Flake"
date: "2017-07-16 07:49:35"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---


```{r echo=TRUE, message=TRUE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE)
library(Rcmdr)
library(car)
library(RcmdrMisc)
library(mice)
library(data.table)
library(chron)
library(ggplot2)
```


```{r echo=TRUE, message=TRUE}
# include this code chunk as-is to enable 3D graphs
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```
1) Code for reading in the dataset and/or processing the data
*Download data zip file from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
*Extract activity data file to local drive
*Run the following code to load data


```{r echo=TRUE, message=TRUE}
activity <- 
  read.table("C:/Users/Rick/Documents/Coursera/Reproducible Research/activity.csv",
   header=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)
```
2) Histogram of the total number of steps taken each day

```{r echo=TRUE, message=TRUE}
activity <- data.table(activity)
setkey(activity,date)
dactivity <- as.data.frame(activity[, sum(steps, na.rm = TRUE),by = date])
names(dactivity)[c(2)] <- c("steps")
with(dactivity, Hist(steps, scale="frequency", breaks="Sturges", col="darkgray"))
```
3) Mean and median number of steps taken each day

```{r echo=TRUE, message=TRUE}
meanday <- with(dactivity,mean(steps))
medianday <- with(dactivity,median(steps))
mactivity <- as.data.frame(activity[, mean(steps, na.rm = TRUE),by = date])
mdactivity <- as.data.frame(activity[, median(steps, na.rm = TRUE),by = date])
names(mactivity)[c(2)] <- c("meansteps")
names(mdactivity)[c(2)] <- c("mediansteps")
stactivity <-merge(mactivity,mdactivity,all.x = TRUE)

```
Mean of days is `r meanday` and Median of days is `r medianday`.


4) Time series plot of the average number of steps taken

```{r echo=TRUE, message=TRUE}
with(activity, plotMeans(steps, date, error.bars="none", connect=TRUE))
```

5) The 5-minute interval that, on average, contains the maximum number of steps

```{r echo=TRUE, message=TRUE}
setkey(activity,interval)
iactivity <- as.data.frame(activity[, mean(steps, na.rm = TRUE),by = interval])
names(iactivity)[c(2)] <- c("steps")
subset(iactivity,steps == max(steps))
```

6) Code to describe and show a strategy for imputing missing data

```{r echo=TRUE, message=TRUE}
# calculating imputed data. It uses linear reqression against the existing values
library(mice)
imactivity <- mice(data = activity, m = 5, method = "pmm", maxit = 5, seed = 500)
#completing data using the 2nd of 5 result sets
completeactivity <- complete(imactivity,2)

```

7) Histogram of the total number of steps taken each day after missing values are imputed

```{r echo=TRUE, message=TRUE}
completeactivity <- data.table(completeactivity)
setkey(completeactivity,date)
dcompactivity <- as.data.frame(completeactivity[, sum(steps, na.rm = TRUE),by = date])
names(dcompactivity)[c(2)] <- c("steps")
with(dcompactivity, Hist(steps, scale="frequency", breaks="Sturges", col="darkgray"))
```

8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends. 


```{r}
library(chron)
completeactivity1 <- completeactivity[,daynum:=wday(date)]
completeactivity1[,weekend:=is.weekend(date)]
 Ciactivity <- as.data.frame(completeactivity1[, mean(steps, na.rm = TRUE),keyby = .(interval,weekend)])
names(Ciactivity)[c(3)] <- c("meansteps")
Ciactivity <- data.table(Ciactivity)
Ciactivity$weekend <- replace(Ciactivity$weekend,Ciactivity$weekend =='TRUE','Weekend')
Ciactivity$weekend <- replace(Ciactivity$weekend,Ciactivity$weekend =='FALSE','Weekday')
library(lattice)
xyplot(data=Ciactivity, meansteps~interval|weekend, type ='l',layout = c(1,2))
```


